### 2008-10-08 trunk/Papers/rmrs-08/rmrs.tex
done with sect 3 for now

In the context of robust semantic processing, this is a desirable
feature of \rmrs, because it means that when we enrich an \rmrs\
obtained from a shallow processor with more semantic information, we
don't change the set of models; we only \emph{restrict} the set of
models further and further towards the semantic representation we are
trying to reconstruct.  Furthermore, it has been shown in the
literature that a dominance-constraint style semantics for
underspecified representations gives us more room to maneuver when
developing efficient solvers than an \mrs-style semantics
\cite{Althaus_etal:JoA}.

@@

Let's look at an example for all these definitions.  All the \rmrs s
presented in Section~\ref{sec:motivation}, where we replace $\qeq$ by
$\dom$, are in solved form; this is least obvious for
(\ref{ex:cat-erg}), but becomes clear once we notice that no label is
on the right-hand side of two dominance atoms.  However, this is
because the model that we can read off of (\ref{ex:cat-erg}) is
similar to the one shown in Fig.~\ref{fig:fat-black-cat}; we simply
take the three subtrees that are specified by ``every fat cat'',
``some dog'', and ``chased'' and combine them as sisters.

@@

So far we have included the dominance relation from
\cite{egg:etal:2001} in the syntax and semantics of the formalism and
not the qeq relation from \cite{copestake:etal:2005}.  This is partly
because dominance is more expressive than qeq; for instance if a
dependency parser links a determiner to a noun which in turn is
linked to a verb, then the dominance relation is required to express
the fact that
that the argument to the predicate that's introduced by the verb is bound by
the quantifier introduced by 
the determiner (see earlier discussion).  However, it is very
straightforward to add qeq to the syntax of the language and to extend
the model theory accordingly.  This extension adds a new atom $X=_q Y$
to Def.~\ref{defn:rmrs-syntax}, and $\tau,\alpha,g,\sigma$ will
satisfy $X=_q Y$ iff $\alpha(X)$ and $\alpha(Y)$ are separated by
zero or more nodes in $\tau$, each of which has a third child that is a
quantifier.  It is possible to replace dominance by qeq in the above
definition of \rmrs\ as well; and all the above propositions about
solved forms still hold if ``dominance'' is systematically replaced
with ``qeq''.

Furthermore, grammar developers such as those in the {\sc delphin}
community typically adopt conventions that restrict them to a
fragment of the language from Def.~\ref{defn:rmrs-syntax} (once
qeq is added to it), or they restrict attention to only a subset of
the models (e.g., any variable that equals the variable
that's bound by a quantifier is also bound by that quantifier). 
Our formalism provides a general framework
into which all these various fragments fit, and it's a matter of
future work to investigate how these fragments afford efficient
algorithms for performing inference over the \rmrs s.

Another feature of the existing \rmrs\ literature is that each term of
an \rmrs\ is equipped 
with a \emph{sort}.  For instance, individual variables are typically
assigned the sort $x$, event variables the sort $e$, and holes the
sort $h$; sometimes subsorts of these are considered (e.g.\ $e_{PAST}$
for events in the past tense), and all the sorts arranged in a sort
hierarchy ${\cal S}$.
While we have defined \rmrs\ without sorts to simplify the
presentation, it is straightforward to add sorts to the definition.
For this, one assumes that the signature $\Sigma$ is sorted, i.e.\
assigns a sort $s_1\times\ldots s_n\rightarrow s$ to each constructor,
where $n$ is the constructor's arity (possibly zero) and
$s,s_1,\ldots,s_n\in {\cal S}$ are atomic sorts.  We can then
restrict the models of \rmrs\ to trees that are \emph{well-sorted} in
the usual sense, i.e.\ those in which we can infer a sort for each
subtree, and require that the variable assignment functions likewise
respect the sorts.  If we then modify Def.~\ref{defn:solved-forms}
such that the constructor $p$ of sufficiently high arity is also
consistent with the sorts of the known arguments---i.e., if $p$ has
sort $s_1 \times \ldots \times s_n \rightarrow s$ and the \rmrs\
contains an atom $\ARG_{\{k\}}(Y,i)$ and $i$ is of sort $s'$, then
$s'$ is a subsort of $s_k$---all the above propositions about solved
forms remain true.

>>>>

In the context of robust semantic processing, this is a desirable
feature of \rmrs, because it means that when we enrich an \rmrs\
obtained from a shallow processor with more semantic information --
such as the relational information introduced by syntactic
constructions such as appositives, noun-noun compounds, and
nonrestrictive relative clauses --, we don't change the set of models;
we only \emph{restrict} the set of models further and further towards
the semantic representation we are trying to reconstruct.
Furthermore, it has been shown in the literature that a
dominance-constraint style semantics for underspecified
representations gives us more room to maneuver when developing
efficient solvers than an \mrs-style semantics
\cite{Althaus_etal:JoA}.

@@

Let's look at an example for all these definitions.  All the \rmrs s
presented in Section~\ref{sec:motivation} (replacing $\qeq$ by
$\dom$), are in solved form; this is least obvious for
(\ref{ex:cat-erg}), but becomes clear once we notice that no label is
on the right-hand side of two dominance atoms.  However, this is
because the model that we can read off of (\ref{ex:cat-erg}) is
similar to the one shown in Fig.~\ref{fig:fat-black-cat}; this model
is problematic in several ways and in particular contains unbound
variables.  If we restrict the class of models we consider to those in
which variables are bound, we can add the information that the
quantifiers outscope their bound variables to the \rmrs\ without
changing its models further -- i.e., we add the atoms $h_3 \dom
l_5$ and $h_8 \dom l_5$.  The extended \rmrs\ is no longer in solved
form because the label $l_5$ is on the right-hand side of two
dominance edges.  Instead, it has the following two solved forms:

\begin{examples}
\item \label{ex:sf-1}
  $l_1\handel a_1\handel\sempred{\_every\_q\_1}(x_1)$,
  $\mathsf{RSTR}(a_1,h_2)$, $\mathsf{BODY}(a_1,h_3)$,
  $l_{41}\handel a_{41}\handel\sempred{\_fat\_j\_1}(e'), \ARG_1(a_{41},x_1)$,
  $l_{41}\handel a_{42}\handel\sempred{\_cat\_n\_1}(x_1)$,
  %
  $l_6\handel a_6\handel\sempred{\_some\_q\_1}(x_6)$,
  $\mathsf{RSTR}(a_6,h_7)$, $\mathsf{BODY}(a_6,h_8)$,
  $l_9\handel a_9\handel\sempred{\_dog\_n\_1}(x_6)$,
  %
  $l_5\handel a_5\handel\sempred{\_chase\_v\_1}(e)$,
  $\ARG_1(a_5,x_1)$, $\ARG_2(a_5,x_6)$,
  $h_2 \dom l_{41}$, $h_3 \dom l_6$,
  $h_7 \dom l_9$, $h_8 \dom l_5$
\item \label{ex:sf-2}
  $l_1\handel a_1\handel\sempred{\_every\_q\_1}(x_1)$,
  $\mathsf{RSTR}(a_1,h_2)$, $\mathsf{BODY}(a_1,h_3)$,
  $l_{41}\handel a_{41}\handel\sempred{\_fat\_j\_1}(e'), \ARG_1(a_{41},x_1)$,
  $l_{41}\handel a_{42}\handel\sempred{\_cat\_n\_1}(x_1)$,
  %
  $l_6\handel a_6\handel\sempred{\_some\_q\_1}(x_6)$,
  $\mathsf{RSTR}(a_6,h_7)$, $\mathsf{BODY}(a_6,h_8)$,
  $l_9\handel a_9\handel\sempred{\_dog\_n\_1}(x_6)$,
  %
  $l_5\handel a_5\handel\sempred{\_chase\_v\_1}(e)$,
  $\ARG_1(a_5,x_1)$, $\ARG_2(a_5,x_6)$,
  $h_2 \dom l_{41}$, $h_3 \dom l_5$,
  $h_7 \dom l_9$, $h_8 \dom l_1$
\end{examples}

Notice that we have eliminated all equalities by unifying the variable
names, and we have fixed the relative scope of the two quantifiers.
Each of these solved forms now stands for a separate class of models;
for instance, the first model in Fig.~\ref{fig:1} is a model of
(\ref{ex:sf-1}), whereas the second is a model of (\ref{ex:sf-2}).

@@

So far we have based the syntax and semantics of \rmrs\ on the
dominance relation from \newcite{egg:etal:2001} rather than the qeq
relation from \newcite{copestake:etal:2005}.  This is partly because
dominance is a weaker relation than qeq: If a dependency parser links
a determiner to a noun and this noun to a verb, then we can use
dominance but not qeq to represent that the predicate introduced by
the verb is outscoped by the quantifier introduced by the determiner
(see earlier discussion).  However, it is very straightforward to add
qeq to the syntax of the language and to extend the model theory
accordingly.  This extension adds a new atom $X=_q Y$ to
Def.~\ref{defn:rmrs-syntax}, and $\tau,\alpha,g,\sigma$ will satisfy
$X=_q Y$ iff $\alpha(X)$ and $\alpha(Y)$ are separated by zero or more
nodes in $\tau$, each of which has a third child that is a
quantifier. All the above propositions about solved forms still hold
if ``dominance'' is systematically replaced with ``qeq''.

Furthermore, grammar developers such as those in the {\sc delph-in}
community typically adopt conventions that restrict them to a fragment
of the language from Def.~\ref{defn:rmrs-syntax} (once qeq is added to
it), or they restrict attention to only a subset of the models (e.g.,
ones in which all variables are bound). Our formalism provides a
general framework into which all these various fragments fit, and it's
a matter of future work to explore these fragments further.

Another feature of the existing \rmrs\ literature is that each term of
an \rmrs\ is equipped with a \emph{sort}.  For instance, individual
variables are typically assigned the sort $x$, event variables the
sort $e$, and holes the sort $h$; these are arranged together with
their subsorts in a sort hierarchy ${\cal S}$.  While we have defined
\rmrs\ without sorts to simplify the presentation, it is
straightforward to add them.  For this, one assumes that the signature
$\Sigma$ is sorted, i.e.\ assigns a sort $s_1\times\ldots
s_n\rightarrow s$ to each constructor, where $n$ is the constructor's
arity (possibly zero) and $s,s_1,\ldots,s_n\in {\cal S}$ are atomic
sorts.  We restrict the models of \rmrs\ to trees that are
\emph{well-sorted} in the usual sense, i.e.\ those in which we can
infer a sort for each subtree, and require that the variable
assignment functions likewise respect the sorts.  If we then modify
Def.~\ref{defn:solved-forms} such that the constructor $p$ of
sufficiently high arity is also consistent with the sorts of the known
arguments---i.e., if $p$ has sort $s_1 \times \ldots \times s_n
\rightarrow s$ and the \rmrs\ contains an atom $\ARG_{\{k\}}(Y,i)$ and
$i$ is of sort $s'$, then $s'$ is a subsort of $s_k$---all the above
propositions about solved forms remain true.
